Inference for Trans-dimensional Bayesian Models with Diffusive Nested Sampling

### 2014-08-27 Documents/Paper/ms.tex
More editing

Some authors have addressed these kinds of problems by
doing separate runs of Nested Sampling
(or another method),
with $N$ fixed at various trial values \citep[e.g.][]{fengji, feroz}.
Then, the posterior for $N$ can be calculated
based on the estimates of the evidence obtained as a function of $N$. However,
in this paper, we use trans-dimensional MCMC moves to infer $N$. In other words,
$N$ is treated as just another parameter, and the marginal likelihood defined
in Equation~\ref{eq:evidence}, which we can compute, also includes the sum over
the unknown parameter $N$. Our motivation
for using Nested Sampling is not primarily to calculate $\mathcal{Z}$.
The motivation for using Nested Sampling, rather than just sampling the
posterior using trans-dimensional MCMC, is that the posterior often contains
difficult features, such as strong dependencies or multiple modes, that cause
problems with mixing. Diffusive Nested Sampling
(DNS) replaces the posterior distribution with an alternative target
distribution composed of a mixture of the prior distribution with other, more
constrained versions of the prior, facilitating mixing between the multiple modes
and along degeneracy curves.
This is similar to the idea of annealed importance sampling
\citep{neal} or parallel tempering \citep{pt},
where the target distribution is modified from the posterior to something
``easier''. However,
the sequence of distributions used in Nested Sampling avoids some of the
problems with annealing-based methods. In particular, there is no need to choose
a temperature schedule (a potentially large set of tuning parameters),
and the method does not fail when phase transitions
are present.

@@

>>>>

Some authors have addressed these kinds of problems by
doing separate runs of Nested Sampling
(or another method),
with $N$ fixed at various trial values \citep[e.g.][]{fengji, feroz}.
Then, the posterior for $N$ can be calculated
based on the estimates of the evidence obtained as a function of $N$.
This approach does not generalise well to large $N$, as it would
require a large number of parallel runs.

In this paper, we introduce an approach that uses trans-dimensional MCMC moves
to infer $N$, but embeds this process in the Nested Sampling framework to
overcome difficulties that would be encountered by sampling the posterior
distribution directly. Our motivation
for using Nested Sampling is not primarily to calculate the marginal
likelihood $\mathcal{Z}$ (including summing over $N$, as in
Equation~\ref{eq:evidence}),
although $\mathcal{Z}$ is readily available from the output.
The motivation for using Nested Sampling, rather than just sampling the
posterior using trans-dimensional MCMC, is that the posterior often contains
difficult features, such as strong dependencies or multiple modes, that cause
problems with mixing. Diffusive Nested Sampling
(DNS) replaces the posterior distribution with an alternative target
distribution composed of a mixture of the prior distribution with other, more
constrained versions of the prior, facilitating mixing between the multiple modes
and along degeneracy curves.
This is similar to the idea of annealed importance sampling
\citep{neal} or parallel tempering \citep{pt},
where the target distribution is modified from the posterior to something
``easier''. However,
the sequence of distributions used in Nested Sampling avoids some of the
problems with annealing-based methods. In particular, there is no need to choose
a temperature schedule (a potentially large set of tuning parameters),
and the method does not fail when phase transitions
are present.

@@

A C++ implementation of the ideas described in this paper is available online
(under the terms of the GNU General Public Licence) at
{\tt https://github.com/eggplantbren/RJObject}.
